---
title: "data_cleaning"
author: "nam"
date: "2025-04-24"
output: html_document
---

# -----------------------------------------------------------------------------
# read files and summarise reports

# read a list of csv files
```{r read_files}
source("customized_functions/read_files_by_path.R")
path <- "../data/merged/"
tables <- read_files_by_path(directory_path = path)
```

# get metadata as a csv file
```{r}
source("customized_functions/get_metadata.R")
metadata_table <- get_metadata(tables)
write.csv(metadata_table, "../metadata/metadata.csv", row.names=FALSE)
```
# skim reports ("before")
```{r}
source("customized_functions/write_skim_report.R")
out_path <- "metadata/skim_summary.csv"
write_skim_report(tables, out_path)
```
# By reading metadata report and skim report:
# -    Missing Data -> Detection: one column in weight_NA table
# -    Duplicates -> No Detection
# -    Extra white spaces in character cols -> No Detection
# -    Column Data Type Issues -> Detection: All date time columns are character
# -    Invalid Values -> No Detection


# -----------------------------------------------------------------------------
# 1.    Missing Data
# weight_NA
```{r}
query <- tables[["weight_NA"]] %>% # count by grouped Id
  group_by(Id, Fat) %>%
  # Count the number of rows in each group.
  # The 'n()' function counts rows within the current group.
  summarise(count = n(), .groups = 'drop') %>% # .groups = 'drop' removes grouping after summarizing
  # Arrange the results by the 'id' column
  arrange(Id)
print(query) 
```
# The column Fat has 94% values missing. 
# Do not have enough values for meaningful filling.
# Drop column Fat
```{r}
tables$weight_NA <- tables$weight_NA %>% select(-Fat) # drop NA
```


# -----------------------------------------------------------------------------
# 2.    Column Data Type (cast datetime columns to the correct datetime types)
```{r include=FALSE}
library(tidyverse)
library(lubridate)
```
# activity_daily
```{r}
tables[["activity_daily"]] %>% head(1) # run this twice (before & after casting) for checking 
```
```{r}
tables$activity_daily$ActivityDate <- mdy(tables$activity_daily$ActivityDate)
```

# calories_hourly
```{r}
tables[["calories_hourly"]] %>% head(10) # run this twice (before & after casting) for checking 
```
```{r}
tables$calories_hourly$ActivityHour <- mdy_hms(tables$calories_hourly$ActivityHour)
```


# calories_minute
```{r}
tables[["calories_minute"]] %>% head(5) # run this twice (before & after casting) for checking 
```
```{r}
tables$calories_minute$ActivityMinute <- mdy_hms(tables$calories_minute$ActivityMinute)
```

# heartrate_second
```{r}
tables[["heartrate_second"]] %>% head(5) # run this twice (before & after casting) for checking 
```
```{r}
tables$heartrate_second$Time <- mdy_hms(tables$heartrate_second$Time)
```

# intensities_hourly
```{r}
tables[["intensities_hourly"]] %>% head(5) # run this twice (before & after casting) for checking 
```
```{r}
tables$intensities_hourly$ActivityHour <- mdy_hms(tables$intensities_hourly$ActivityHour)
```

# intensities_minute
```{r}
tables[["intensities_minute"]] %>% head(5) # run this twice (before & after casting) for checking 
```
```{r}
tables$intensities_minute$ActivityMinute <- mdy_hms(tables$intensities_minute$ActivityMinute)
```

# mets_minute
```{r}
tables[["mets_minute"]] %>% head(5) # run this twice (before & after casting) for checking 
```
```{r}
tables$mets_minute$ActivityMinute <- mdy_hms(tables$mets_minute$ActivityMinute)
```

# sleep_minute
```{r}
tables[["sleep_minute"]] %>% head(5) # run this twice (before & after casting) for checking 
```
```{r}
tables$sleep_minute$date <- mdy_hms(tables$sleep_minute$date)
```

# steps_hourly
```{r}
tables[["steps_hourly"]] %>% head(5) # run this twice (before & after casting) for checking 
```
```{r}
tables$steps_hourly$ActivityHour <- mdy_hms(tables$steps_hourly$ActivityHour)
```

# steps_minute
```{r}
tables[["steps_minute"]] %>% head(5) # run this twice (before & after casting) for checking 
```
```{r}
tables$steps_minute$ActivityMinute <- mdy_hms(tables$steps_minute$ActivityMinute)
```

# weight_NA
```{r}
tables[["weight_NA"]] %>% head(5) # run this twice (before & after casting) for checking 
```
```{r}
tables$weight_NA$Date <- mdy_hms(tables$weight_NA$Date)
```

# -----------------------------------------------------------------------------
# Write cleaned tables to csv files
```{r}
source("./customized_functions/write_files_to_csv.R")
write_files_to_csv(list_of_tables=tables, path="../data/cleaned")
```

# skim reports ("after")
```{r}
source("customized_functions/write_skim_report.R")
out_path <- "metadata/skim_summary_after.csv"
write_skim_report(tables, out_path)
```