---
title: "file_ingestion_to_msql"
output: html_notebook
---

# --- R Notebook Chunk 1: Load Libraries ---
# Install if you don't have them:
``` {r install_dependencies}
install.packages("readr")
install.packages("RMariaDB")
library(readr)
library(RMariaDB)
```

# --- R Notebook Chunk 2: Database Connection Setup ---
# Replace with your actual database details
```{r database_config}
db_user <- "root"
db_password <- "admin"
db_name <- "fitbit_fitness_tracker_data"
db_host <- "localhost" # Or your server IP/hostname
db_port <- 3306        # Default MySQL port
```

# Establish the connection
``` {r connect_to_db}
con <- dbConnect(RMariaDB::MariaDB(),
                 user = db_user,
                 password = db_password,
                 dbname = db_name,
                 host = db_host,
                 port = db_port)
```

# Check if connection is successful
``` {r validate_db_connection}
if (dbIsValid(con)) {
  message("Successfully connected to the database.")
} else {
  stop("Database connection failed!")
}
```

# --- R Notebook Chunk 3: Ingestion Function ---
```{r ingestion_function}
ingest_csv_to_mysql <- function(file_path, db_connection, table_name, ...) {
#' Ingests a CSV file into a MySQL database table.
#'
#' @param file_path The full path to the CSV file.
#' @param db_connection A valid database connection object.
#' @param table_name The name for the new database table.
#' @param ... Additional arguments passed to readr::read_csv and dbWriteTable.
#' @return TRUE if successful, FALSE otherwise.

# File Existence Check 
   if (!file.exists(file_path)) {
    warning("File not found: ", file_path)
    return(FALSE)
   }
  
# Reading the CSV File with Error Handling
  message("Reading file: ", file_path)
  # Use read_csv for robust parsing
  data_to_ingest <- tryCatch({
    readr::read_csv(file_path, show_col_types = FALSE, ...)
  }, error = function(e) {
    warning("Error reading CSV ", file_path, ": ", e$message)
    return(NULL)
  })
  
# Handling CSV Read Failure (Continued)
  if (is.null(data_to_ingest)) {
    return(FALSE)
  }
  
# Writing Data to the Database with Error Handling
  message("Writing data to MySQL table: ", table_name)
  # Write data to the database table
  # Use overwrite = TRUE to replace the table if it exists
  # Use append = TRUE to add rows to an existing table
  success <- tryCatch({
    dbWriteTable(db_connection, name = table_name, value = data_to_ingest, overwrite = TRUE, ...) # Choose overwrite or append
    TRUE
  }, error = function(e) {
    warning("Error writing to DB table ", table_name, ": ", e$message)
    return(FALSE)
  })

  
# Reporting Success  
  if (success) {
    message("Successfully ingested ", file_path, " into table ", table_name)
  }
  
# Returning the Final Status
  return(success)
}
```

# --- R Notebook Chunk 4: Ingest all CSVs in WD ---
```{r} 
# List all files ending with .csv in the current working directory (.)
csv_files_in_wd <- list.files(path = ".", pattern = "\\.csv$", full.names = TRUE)

if (length(csv_files_in_wd) == 0) {
  message("No CSV files found in the working directory.")
} else {
  message("Found ", length(csv_files_in_wd), " CSV files.")
  for (file in csv_files_in_wd) {
    # Create a table name from the file name (e.g., "my_data.csv" -> "my_data")
    table_name <- tools::file_path_sans_ext(basename(file))
    # Ensure valid MySQL table names (e.g., replace hyphens with underscores if needed)
    table_name <- gsub("-", "_", table_name)

    # Call the ingestion function for each file
    ingest_csv_to_mysql(file_path = file, db_connection = con, table_name = table_name)
  }
}
```




# --- R Notebook Chunk 4: Example Usage (Ingest all CSVs in WD) ---

# Set working directory
```{r set_wd}
setwd("C:/Users/user/Desktop/GG DRIVE/DA/Course_DA_Google/case_study/bella/FitBit Fitness Tracker Data")
```

``` {r}
# List all files ending with .csv in the current working directory (.)
csv_files_in_wd <- list.files(path = ".", pattern = "\\.csv$", full.names = TRUE)

# ingest and create table in db using file names
if (length(csv_files_in_wd) == 0) {
  message("No CSV files found in the working directory.")
} else {
  message("Found ", length(csv_files_in_wd), " CSV files.")
  for (file in csv_files_in_wd) {
    # Create a table name from the file name (e.g., "my_data.csv" -> "my_data")
    table_name <- tools::file_path_sans_ext(basename(file))
    # Ensure valid MySQL table names (e.g., replace hyphens with underscores if needed)
    table_name <- gsub("-", "_", table_name)
    # Call the ingestion function for each file
    ingest_csv_to_mysql(file_path = file, db_connection = con, table_name = table_name)
  }
}
```

# --- R Notebook Chunk 5: Close Database Connection ---
# Good practice to close the connection when done
```{r}
if (dbIsValid(con)) {
  dbDisconnect(con)
  message("Database connection closed.")
}
```